Mechanical loading prediction through accelerometry data during walking and running

### 2021-09-24 manuscript/manuscript.Rmd
Add new study from Florencio in the discussion

The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate mechanical loading during walking and running. Predictions were tested with accelerometers placed at different body locations, using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running. Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all of the pGRF prediction models developed presented a MAPE below the 8\% threshold. However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.

@@

Besides these advantages, our models present some important potential clinical applications. For instance, these models can be used to analyse data from large observational studies and make associations between mechanical loading and bone health. But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation. Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps. This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or also in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].

@@

In conclusion, pGRF could be accurately predicted by accelerometry-based regression models. These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip. The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanican loading and either beneficial or pathological osteoarticular adaptations. This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was unsatisfactory.

>>>>

The purpose of this study was to develop pGRF and pLR prediction equations based on accelerometer data able to accurately estimate mechanical loading during walking and running. Predictions were tested with accelerometers placed at different body locations, using both the resultant vector and its vertical component in a set of incremental speeds, ranging from slow walking to running. Our findings demonstrated that accelerometer-based regression models were a valid and accurate method to estimate pGRF during human locomotion, as all pGRF prediction models developed presented a MAPE below the 8\% threshold. However, accelerometry-based equations developed to predict pLR did not yield trustworthy results.

@@

Besides these advantages, our models present some important potential clinical applications. For instance, these models can be used to analyse data from large observational studies and make associations between mechanical loading and bone health. But instead of analysing accelerometer data averaged in epochs similarly to some previous publications [@Chastin2014; @Stiles2017], which can dilute the higher intensities [@Fabre2020], with our method raw ACC can be analysed at a higher resolution by identifying the pGRF associated with each step taken and establishing, for example, the number of peaks above a certain magnitude threshold needed to elicit a positive bone adaptation. Another potential application could be the computation of the cumulative mechanical loading by adding up the pGRF of all detected steps [@Diniz-Sousa2021]. This information could be used either for the determination of the relationship between the mechanical loading accumulated in a certain period (e.g., a day) and the bone response to this loading [@Ahola2010], or to investigate the causes of bone loss in populations with bone health issues, such as bariatric patients [@Diniz-Sousa2021]. This approach has already been applied with equations developed in a previous research from our group [@Veras2020], but now can be extended to include both types of locomotion (walking and running) contemplated in the equations from the current study. Another possible application is in sports training, by monitoring the amount of loading beared by athletes and relate this information with osteoarticular safety of the training regimen [@Vanwanseele2020].

@@

In conclusion, pGRF could be accurately predicted by accelerometry-based regression models. These models are valid for the use of both the resultant vector and its vertical component, and using data from accelerometers placed at either the ankle, lower back or hip. The prediction models developed in this study have proved to be an efficient method to assess pGRF out of laboratory conditions, with potential applications for monitoring the mechanical loading in several contexts and establishing a relationship between mechanical loading and either beneficial or pathological osteoarticular adaptations. This is also the first work, to the best of our knowledge, to explore the prediction of pLR based on accelerometry data during both walking and running, although the accuracy of these models was unsatisfactory.
